\chapter{Evaluation}
\label{chap:evaluation}

The ticketing system developed during this project successfully implements the core functionality required for submitting and managing support requests. Users can create tickets through a structured interface, while the system demonstrates integration with modern web technologies and an AI component designed to assist with handling queries. The implementation meets the primary objectives established to deliver a functional Minimum Viable Product (MVP) within the project timeframe. Despite this, several limitations remain that affect the system’s current capabilities and highlight opportunities for further development.

\section{Limitations}
One limitation relates to the integration of artificial intelligence within the system. The current implementation relies on a Gemini API key, which is provided through a free student-tier account. While this allows the AI functionality to operate, it introduces usage limitations and may not be suitable for large-scale deployment. Additionally, the API key currently functions reliably only within the deployed version of the system, which restricts the ability to test certain AI features during local development.

Another limitation concerns the system’s knowledge support features. Although an FAQ section is included, the process for expanding and managing frequently asked questions could be improved. A more flexible interface that allows administrators to easily add or modify questions would make the system more adaptable to evolving user needs.

Certain integrations that could improve usability are also not yet implemented. For example, the system does not currently support ticket creation through email, which would allow students to automatically generate support requests by sending an email. Similarly, integration with university platforms has not been fully explored. Connecting the system to institutional services such as Microsoft 365 could enable additional features, including automatically scheduling meetings with support staff through Microsoft Teams.

Another objective that the team had aimed to achieve was enabling staff to address multiple tickets regarding the same query collectively, rather than handling each ticket individually. This feature would have streamlined responses and reduced duplicate effort. However, due to time constraints and the approaching project deadline, the team chose to prioritise completing the core functionality and ensuring the quality of the implemented features. As a result, this objective was deferred, highlighting an area for improvement in future development.

\section{Future Possibilites}
There are several opportunities to extend the system in future development. One important area is scalability. The current architecture could be expanded to support a larger number of users by scaling both the database infrastructure and the deployed application environment. This would allow the system to handle a significantly higher volume of requests while maintaining performance.

Another promising direction is the expanded use of artificial intelligence within the support process. Rather than only assisting after a ticket has been created, the AI component could analyse a user’s question as it is typed and proactively suggest relevant answers or guide users to specific sections of existing documentation that directly address their issue. This approach would reduce unnecessary ticket submissions and improve efficiency for both users and support staff.

Several constraints affected what could be delivered within the project scope and environment.

\textbf{Gemini API credentials and deployment.} The AI assistant relies on the Google Gemini API, which requires a secret API key. For security, the key cannot be distributed to every developer machine; as a result, the AI feature is only fully exercised in the deployed environment where credentials are configured securely. Local development therefore does not mirror production behaviour for this component.

\textbf{API usage limits.} The team used a student-oriented Gemini offering with a free usage tier. Daily request and credit limits apply, so sustained load-testing or heavy classroom demos could exhaust the quota. This caps how aggressively the assistant can be used in evaluation and constrains real-world adoption without a paid or institutional plan.

\textbf{Ticket priority defaults.} Support tickets are not assigned fine-grained priorities in the current implementation; a medium default is applied broadly because explicit priority workflows were not essential to the core ticketing and meeting features. This simplifies the model but does not reflect nuanced urgency in a production helpdesk.

\textbf{FAQ content management.} A feature allowing users to add new FAQ entries was not implemented. Besides time constraints, the team judged that an open submission form could clutter the interface and raised unresolved policy questions: whether students, staff, or administrators should author or approve changes. A minimal, read-only FAQ list was preferred for consistency and moderation control.

\textbf{Email notifications.} Automated email (for example via Mailgun or similar APIs) was explored in a prototype but abandoned. Third-party transactional email typically incurs cost, and the integration proved brittle relative to project time. The final system does not rely on outbound email for ticket updates.

\textbf{University integrations and realistic staff data.} Deeper integration was considered, including Microsoft 365--based flows (for example Teams meeting links) and live staff calendars or working hours from institutional systems. The university could not provide real-time APIs or feeds for staff availability in a form suitable for this project. Designing and securing such integrations would also have added substantial complexity. The team therefore used seeded, representative office hours rather than live directory data, which limits fidelity to real departmental schedules.

\section{Future possibilities}

\textbf{Scalability and infrastructure.} The production database is hosted on Neon. The free tier caps storage (for example around 0.5\,GB); the project currently uses a small fraction of that (approximately 0.03\,GB), so there is headroom for growth. Neon’s paid tiers offer higher limits and can support a larger dataset and connection load. The deployed application stack can be scaled horizontally or vertically as traffic grows, and the Gemini integration can be upgraded to higher quotas or enterprise endpoints if AI usage becomes business-critical.

\textbf{AI-first resolution before ticketing.} A natural extension of the current Gemini-powered helper is to deepen its role in the support funnel: students would interact with the assistant first for routine questions, with clear escalation paths only when the model is uncertain or the issue requires human handling. This would reduce duplicate tickets, align with the design goal of first-line automation, and could be combined with analytics on which queries resolve without a ticket. Future work could include stronger grounding in official KCL documentation, explicit confidence thresholds for escalation, and tighter coupling between assistant sessions and ticket creation so context carries over when a formal request is needed.
